A Domain Specific Language for SpectreGuard

### 2020-12-11 paper/main.tex
wibble

  A Spectre mitigation strategy in the form of a mechanism for marking variables
  as ``non-speculative'' was developed in \cite{SpectreGuard}. In this paper, we
  develop an embedded domain specific language (EDSL) in Haskell to provide
  support for this feature, with a backend that generates C code. The static
  type system of Haskell is used so that many basic forms of copying secret
  variables to public variables are disallowed. Additionally, a static analysis
  was developed which will detect some cases of secret variables leaking to
  public variables via control flow constructs.

@@

\section{Introduction}
\label{sec:Introduction}
SpectreGuard is provides an abstraction that allows programmers to prevent
speculative access of certain variables.~\cite{SpectreGuard} As Spectre-based
attacks remain an issue, even in modern computer systems, such mitigations and
preventative measures are crucial to maintaining the security of software.  For
preserving program correctness, static analyses are desirable to prevent
programmers from accidentally successfully compiling code which copies a
``secret'' variable (a variable protected by SpectreGuard) to a ``public''
variable (a variable with no such protection). To this end, two interrelated
static analysis techniques are designed and implemented.

@@

Haskell is a purely functional programming language with strong static type
system based on System F and algebraic datatypes (ADTs).~\cite{HistoryOfHaskell}
These features, particularly the type system, the algebraic datatypes and its
support for higher-order functions, make it a powerful tool for the creation of
EDSLs. Expressions, commands and programs (compound commands) can be
straightforwardly expressed using ADTs using a \textit{deep
embedding}.~\cite{DeepAndShallow}

@@

The one major aspect of the C code generator that is not described in this table
is how the loop variable in \ttt{for} loops can be tagged as secret (if appropriate).
This is accomplished by recovering the type-level sensitivity type information at
runtime. Such a recovery is made possible by the use of singletons.~\cite{Singletons}
\ttt{for} loops provide a particular example of where Haskell's support for higher-order function
can be used inside an EDSL to implement commands that involve the creation of
new variables.

@@

Singletons~\cite{Singletons} are used to preserve this type level at runtime, so that the
code generator can generate SpectreGuard annotations where appropriate. These annotations can
then be used by the SpectreGuard mitigation system.~\cite{SpectreGuard}

@@

\section{Future Work}
\label{sec:FutureWork}
A further investigation into the tradeoffs and possible additional features of
the information flow analysis is likely to be useful. This could provide a more
accurate and succinct picture to programmers of where leaks are occurring, or
potentially occurring. It is possible that some of these other leak conditions
could be incorporated into the type system, by exploiting the dependent type
features enabled by singletons~\cite{Singletons} and, in the future, by
Dependent Haskell.~\cite{DepHaskSpec} Dependent types provide a powerful
technique for writing programs at a type-level, which will be executed at
compile-time by the type checker.~\cite{CertProg, DepHaskSpec} A more
lightweight technique, which might be able to be applied to allow for more
expressive policies to control of information flow within the type system, would
be the ``ghosts of departed proofs'' design technique.~\cite{DepartedProofs}

Another possible, related, future development would be the to attempt to add a
mechanism which automatically fixes some (or all) leaks. Such a system has precendent
in the previously mentioned paper \textit{Liquid Information Flow Control}.~\cite{Lifty}
This could give programmers an additional tool towards ensuring correctness, with
respect to SpectreGuard annotations.

Additionally, features to safely convert between sensitivity levels could prove
useful. A conversion from public to secret would be straightforward. A situational conversion
going the other direction could be provided by an instruction in the EDSL
which might perform a Spectre mitigation measure, such as a cache flushes at appropriate times.~\cite{PLtea-james}

@@

Additional primitive operations added to the EDSL could include higher-level language features, such as higher-order
functions, algebraic data types, pattern matching and tail recursion. An implementation path for
these features is described in the upcoming paper \textit{On Adding Pattern Matching to Haskell-based Deeply Embedded Domain Specific Languages}, by
David Young, Mark Grebe and Andy Gill.~\cite{PatMatchingEDSL}

>>>>

  A Spectre mitigation strategy in the form of a mechanism for marking variables
  as ``non-speculative'' was developed in the form of \textit{SpectreGuard}\cite{SpectreGuard}. In this paper, we
  develop an embedded domain specific language (EDSL) in Haskell to provide
  support for this feature, with a backend that generates C code. The static
  type system of Haskell is used so that many basic forms of copying secret
  variables to public variables are disallowed. Additionally, a static analysis
  was developed which will detect some cases of secret variables leaking to
  public variables via control flow constructs.

@@

\section{Introduction}
\label{sec:Introduction}
SpectreGuard is provides an abstraction that allows programmers to prevent
speculative access of certain variables~\cite{SpectreGuard}. As Spectre-based
attacks remain an issue, even in modern computer systems, such mitigations and
preventative measures are crucial to maintaining the security of software.  For
preserving program correctness, static analyses are desirable to prevent
programmers from accidentally successfully compiling code which copies a
``secret'' variable (a variable protected by SpectreGuard) to a ``public''
variable (a variable with no such protection). To this end, two interrelated
static analysis techniques are designed and implemented.

@@

Haskell is a purely functional programming language with strong static type
system based on System F and algebraic datatypes (ADTs)~\cite{HistoryOfHaskell}.
These features, particularly the type system, the algebraic datatypes and its
support for higher-order functions, make it a powerful tool for the creation of
EDSLs. Expressions, commands and programs (compound commands) can be
straightforwardly expressed using ADTs using a \textit{deep
embedding}~\cite{DeepAndShallow}.

@@

The one major aspect of the C code generator that is not described in this table
is how the loop variable in \ttt{for} loops can be tagged as secret (if appropriate).
This is accomplished by recovering the type-level sensitivity type information at
runtime. Such a recovery is made possible by the use of singletons~\cite{Singletons}.
\ttt{for} loops provide a particular example of where Haskell's support for higher-order function
can be used inside an EDSL to implement commands that involve the creation of
new variables.

@@

Singletons~\cite{Singletons} are used to preserve this type level at runtime, so that the
code generator can generate SpectreGuard annotations where appropriate. These annotations can
then be used by the SpectreGuard mitigation system~\cite{SpectreGuard}.

@@

\section{Future Work}
\label{sec:FutureWork}
A further investigation into the tradeoffs and possible additional features of
the information flow analysis is likely to be useful. This could provide a more
accurate and succinct picture to programmers of where leaks are occurring, or
potentially occurring. It is possible that some of these other leak conditions
could be incorporated into the type system, by exploiting the dependent type
features enabled by singletons~\cite{Singletons} and, in the future, by
Dependent Haskell.~\cite{DepHaskSpec} Dependent types provide a powerful
technique for writing programs at a type-level, which will be executed at
compile-time by the type checker~\cite{CertProg, DepHaskSpec}. A more
lightweight technique, which might be able to be applied to allow for more
expressive policies to control of information flow within the type system, would
be the ``ghosts of departed proofs'' design technique~\cite{DepartedProofs}.

Another possible, related, future development would be the to attempt to add a
mechanism which automatically fixes some (or all) leaks. Such a system has precendent
in the previously mentioned paper \textit{Liquid Information Flow Control}~\cite{Lifty}.
This could give programmers an additional tool towards ensuring correctness, with
respect to SpectreGuard annotations.

Additionally, features to safely convert between sensitivity levels could prove
useful. A conversion from public to secret would be straightforward. A situational conversion
going the other direction could be provided by an instruction in the EDSL
which might perform a Spectre mitigation measure, such as a cache flushes at appropriate times~\cite{PLtea-james}.

@@

Additional primitive operations added to the EDSL could include higher-level language features, such as higher-order
functions, algebraic data types, pattern matching and tail recursion. An implementation path for
these features is described in the upcoming paper \textit{On Adding Pattern Matching to Haskell-based Deeply Embedded Domain Specific Languages}, by
David Young, Mark Grebe and Andy Gill~\cite{PatMatchingEDSL}.
